---
title: 'Case Study : Indian Premier League'
author: "Barun Kumar"
date: "12 December 2018"
output: html_document
---

#-------------------------------------------------------------------------
```{r}

library(readxl)
library(dplyr)
#Case Study IPL datasets
#Datasets files : 6 files
ball_by_ball <- read_excel("D:/DSML-M/SEM1/R Language/CSV files/IPL/Ball_by_Ball.xlsx")
match_ipl <- read_excel("D:/DSML-M/SEM1/R Language/CSV files/IPL/Match.xlsx")
player <- read_excel("D:/DSML-M/SEM1/R Language/CSV files/IPL/Player.xlsx")
player_match <- read_excel("D:/DSML-M/SEM1/R Language/CSV files/IPL/Player_Match.xlsx")
season <- read_excel("D:/DSML-M/SEM1/R Language/CSV files/IPL/Season.xlsx")
team <- read_excel("D:/DSML-M/SEM1/R Language/CSV files/IPL/Team.xlsx")

```

#-------------------------------------------------------------------------
1). Top 10 batsman


1a) based on Runs scored.

```{r}

T <- merge(x = ball_by_ball[,c("Striker_Id","Batsman_Scored")], 
          y = player[,c("Player_Id","Player_Name")], 
          by.x = "Striker_Id",by.y = "Player_Id",all.x = TRUE)
 
q1a_top10_run_scored <- T %>% group_by(Player_Name) %>% 
       summarise(Total_Runs = sum(as.integer(Batsman_Scored),na.rm = T)) %>%
       arrange(desc(Total_Runs)) %>% head(10)

q1a_top10_run_scored

```
#-------------------------------------------------------------------------

1b) based on Number of 6s

```{r}

q1b_top10_6s <- T %>% group_by(Player_Name) %>% 
  summarise(Total_6s = sum(as.integer(Batsman_Scored == 6),na.rm = T)) %>%
  arrange(desc(Total_6s)) %>% head(10)

q1b_top10_6s

```

#-------------------------------------------------------------------------

1c) based on Number of 4s  

```{r}

q1c_top10_4s <- T %>% group_by(Player_Name) %>% 
      summarise(Total_4s = sum(as.integer(Batsman_Scored == 4),na.rm = T)) %>%
      arrange(desc(Total_4s)) %>% head(10)

q1c_top10_4s

```

#-------------------------------------------------------------------------

1d) based on Number of 100s

```{r}

T100 <- merge(x = ball_by_ball[,c("Match_Id","Striker_Id","Batsman_Scored")], 
          y = player[,c("Player_Id","Player_Name")], 
          by.x = "Striker_Id",by.y = "Player_Id",all.x = TRUE)


q1d_top10_100s <- T100 %>% group_by(Player_Name,Match_Id) %>% 
   summarise(Total_Runs = sum(as.integer(Batsman_Scored),na.rm = T)) %>%
   filter(Total_Runs > 99) %>% summarise(Total_Centuries = n()) %>%
   arrange(desc(Total_Centuries)) %>% head(10)

q1d_top10_100s

```

#-------------------------------------------------------------------------

2). Top 10 teams

 
2a) Based on number of matches won

```{r}

q2_a_m <- merge(x = match_ipl[,c("Match_Winner_Id")], y = team[,c("Team_Id","Team_Name")],
             by.x = "Match_Winner_Id",by.y = "Team_Id",all.x = TRUE)
q2_a <- q2_a_m %>% group_by(Team_Name) %>% summarise(Match_Won = n()) %>%
      arrange(desc(Match_Won))

q2_a

```

#-------------------------------------------------------------------------

2b) Based on winning the toss and winning the match

```{r}

q2_b_m <- merge(x = match_ipl[,c("Toss_Winner_Id","Match_Winner_Id")], 
                y = team[,c("Team_Id","Team_Name")],
               by.x = "Match_Winner_Id",by.y = "Team_Id",all.x = TRUE)
q2_b <- q2_b_m %>% group_by(Team_Name) %>% filter(Toss_Winner_Id == Match_Winner_Id) %>%
       summarise(Win_Toss_Match = n()) %>% arrange(desc(Win_Toss_Match)) %>% head(10)

q2_b

```


3). Season Wise Holder


3a)	Orange Cap holder 


```{r}

q3_a_m <- merge(x = season[,c("Season_Year","Orange_Cap_Id")], 
                y = player[,c("Player_Id","Player_Name")],
                by.x = "Orange_Cap_Id",by.y = "Player_Id",all.x = TRUE)


q3_a  <- q3_a_m %>% select(Season_Year,Player_Name) %>%
         arrange(Season_Year)

q3_a

```



#3b)	Purple Cap holder 

```{r}

q3_b_m <- merge(x = season[,c("Season_Year","Purple_Cap_Id")], 
                y = player[,c("Player_Id","Player_Name")],
                by.x = "Purple_Cap_Id",by.y = "Player_Id",all.x = TRUE)


q3_b  <- q3_b_m %>% select(Season_Year,Player_Name) %>%
  arrange(Season_Year)

q3_b

```




4). For each season matches with biggest win margin based on runs.


```{r}

q4 <- match_ipl %>% filter(Win_Type == 'by runs') %>% group_by(Season_Id) %>% 
      summarise(Biggest_Win_Margin = max(as.integer(Won_By),na.rm = TRUE)) 
  
  
q4

```



5).	Total Number of : 


5a) Matches


```{r}

q5_a <- match_ipl %>% summarise(no_of_matches = n_distinct(Match_Id))

q5_a

```



5b). 100s


```{r}

q5_b <- T100 %>% group_by(Player_Name,Match_Id) %>% 
  summarise(Total_Runs = sum(as.integer(Batsman_Scored),na.rm = T)) %>%
  filter(Total_Runs > 99) %>% summarise(centuries_total = n()) %>%
  summarise(Total_Centuries = sum(centuries_total))

q5_b

```


5c) 50s

```{r}

q5_c <- T100 %>% group_by(Player_Name,Match_Id) %>% 
  summarise(Total_Runs = sum(as.integer(Batsman_Scored),na.rm = T)) %>%
  filter(Total_Runs > 49) %>% summarise(fifties_total = n()) %>%
  summarise(Total_50s = sum(fifties_total))

q5_c

```


5d) 6s


```{r}

q5_d <- sum(as.integer(ball_by_ball$Batsman_Scored == 6),na.rm = T)

q5_d

```



5e) based on Number of 4s  

```{r}

q5_e <- sum(as.integer(ball_by_ball$Batsman_Scored == 4),na.rm = T)

q5_e

```



6). Team wise count of title win

```{r}

q6 <- match_ipl %>% group_by(Season_Id) %>% filter(Match_Date == max(Match_Date)) %>%
      summarise(Team_Won_Id = Match_Winner_Id) %>% group_by(Team_Won_Id) %>%
      summarise(Title_Win_Count = n())


q6

```

