---
title: 'Case Study : hflights'
author: "Barun Kumar"
output: html_document
---

1.	Install the package "hflights" and load the library command.

```{r}

# install.packages('hflights')

library(hflights)
library(dplyr)
head(hflights)
str(hflights)

```


2.	Convert to a local data frame (use command tbl_df)

```{r}

tflights <- tbl_df(hflights)
tflights

```


3.	Convert to a normal data frame to see all of the columns


```{r}

h <- data.frame(hflights)
head(h)

```


4.	Base R approach & dplyr approach to view all flights on January 1.

```{r}

#Base R approach
tflights[tflights$Month == 1 & tflights$DayofMonth == 1,]

```




```{r}

#dplyr approach            
fl_jan1 <- tflights %>% filter(DayofMonth == 1 & Month == 1)  

fl_jan1

```


5.	Select those flights where UniqueCarrier is "UA" or "AA"

```{r}

#Base R approach
tflights[tflights$UniqueCarrier == "UA" | tflights$UniqueCarrier == "AA",] 
```




```{r}

#dplyr approach 
fl_uniqcar <- tflights %>% filter(UniqueCarrier %in% c("UA", "AA")) 

fl_uniqcar

```


6.	Pick columns by name - DepTime, ArrTime, FlightNum using base r and dplyr.

```{r}

#Base R approach
tflights[,c("DepTime","ArrTime","FlightNum")]

```




```{r}

#dplyr approach
fl_distcol <- tflights %>% select(DepTime,ArrTime,FlightNum)

fl_distcol

```


7.	Select UniqueCarrier and DepDelay columns and filter for delays over 60 minutes in dplyr

```{r}

fl_delay60min <- tflights %>% select(UniqueCarrier,DepDelay) %>% filter(DepDelay > 60) 

fl_delay60min

```


8.	Use mutate to create a new variable Speed (in mph) and  display distance ,airtime and speed columns.

9.	store the data into variable "flights"


```{r}

fl_distcol <- tflights %>% select(Distance,AirTime) %>% 
                            mutate(Speed = 60*Distance/AirTime)
fl_distcol

```


10.	Create a dataset grouped by Dest, and then summarise each group by taking the mean of ArrDelay.

```{r}

f1_grpdel <- tflights %>% group_by(Dest) %>%
             summarise(avg_delay = mean(ArrDelay, na.rm=TRUE))
f1_grpdel

```


11.	For each day of the year, count the total number of flights and sort in descending order.

```{r}

f1_flightc <- tflights %>% group_by(Month, DayofMonth) %>% summarise(flight_count = n()) %>%
              arrange(desc(flight_count))

f1_flightc  

```


12.	For each destination, count the total number of flights and the number of distinct planes that flew there.

```{r}

f1_flightd <- tflights %>% group_by(Dest) %>%
              summarise(flight_count = n(), plane_count = n_distinct(TailNum))

f1_flightd

```


13.	For each destination, show the number of cancelled and not cancelled flights.

```{r}

f1_flightcnc <- tflights %>% group_by(Dest) %>%
                select(Cancelled) %>% table() %>% head()

f1_flightcnc

```

